---
title: "Project 1"
author: "Michael Moerk"
date: "`r Sys.Date()`"
output: pdf_document
---
#Load libraries and data
```{r setup, include=FALSE}
source("http://www.openintro.org/stat/data/cdc.R")
cdc$loss <- cdc$weight - cdc$wtdesire
cdc$gender <- ifelse(cdc$gender == 'm',1,0) #male is 1, female is 0
cdcEdit <- cdc[which(cdc$loss>=0),] #create new dataset for which those who want to gain weight are excluded
library(pROC)
library(xtable)
library(car)
library(statmod)
library(corrplot)
library(ggplot2)
library(dplyr)
```
#Load Emprical Logit plot funtion
```{r}
emplogit = function(x, y, binsize = NULL, ci = FALSE, probit = FALSE, prob = FALSE, main = NULL, xlab = "", ylab = ""){
  # x         vector with values of the independent variable
  # y         vector of binary responses
  # binsize   integer value specifying bin size (optional)
  # ci        logical value indicating whether to plot approximate
  #           confidence intervals (not supported as of 02/08/2015)
  # probit    logical value indicating whether to plot probits instead
  #           of logits
  # prob      logical value indicating whether to plot probabilities
  #           without transforming
  #
  # the rest are the familiar plotting options

  if (length(x) != length(y))
    stop("x and y lengths differ")
  if (any(y < 0 | y > 1))
    stop("y not between 0 and 1")
  if (length(x) < 100 & is.null(binsize))
    stop("Less than 100 observations: specify binsize manually")

  if (is.null(binsize)) binsize = min(round(length(x)/10), 50)

  if (probit){
    link = qnorm
    if (is.null(main)) main = "Empirical probits"
  } else {
    link = function(x) log(x/(1-x))
    if (is.null(main)) main = "Empirical logits"
  }

  sort = order(x)
  x = x[sort]
  y = y[sort]
  a = seq(1, length(x), by=binsize)
  b = c(a[-1] - 1, length(x))

  prob = xmean = ns = rep(0, length(a)) # ns is for CIs
  for (i in 1:length(a)){
    range = (a[i]):(b[i])
    prob[i] = mean(y[range])
    xmean[i] = mean(x[range])
    ns[i] = b[i] - a[i] + 1 # for CI 
  }

  extreme = (prob == 1 | prob == 0)
  prob[prob == 0] = min(prob[!extreme])
  prob[prob == 1] = max(prob[!extreme])

  g = link(prob) # logits (or probits if probit == TRUE)

  linear.fit = lm(g[!extreme] ~ xmean[!extreme])
  b0 = linear.fit$coef[1]
  b1 = linear.fit$coef[2]

  loess.fit = loess(g[!extreme] ~ xmean[!extreme])

    plot(xmean, g, main=main, xlab=xlab, ylab=ylab)
    abline(b0,b1)
    lines(loess.fit$x, loess.fit$fitted, lwd=2, lty=2)
}
```
#Find missing rows
```{r}
missingRows <- which(is.na(cdc))
cdc_levels <- as.factor(cdc$exerany)
levels(cdc_levels)
```

#Generate general statistics about the data
```{r}
cdclevel0 <- cdc[cdc$exerany == 0,]
cdclevel1 <- cdc[cdc$exerany == 1,]
median(cdclevel0$age)
quantile(cdclevel0$age, 0.25)
quantile(cdclevel0$age, 0.75)

median(cdclevel1$age)
quantile(cdclevel1$age, 0.25)
quantile(cdclevel1$age, 0.75)

median(cdclevel0$weight)
quantile(cdclevel0$weight, 0.25)
quantile(cdclevel0$weight, 0.75)

median(cdclevel1$weight)
quantile(cdclevel1$weight, 0.25)
quantile(cdclevel1$weight, 0.75)

median(cdclevel0$height)
quantile(cdclevel0$height, 0.25)
quantile(cdclevel0$height, 0.75)

median(cdclevel1$height)
quantile(cdclevel1$height, 0.25)
quantile(cdclevel1$height, 0.75)

median(cdclevel0$wtdesire)
quantile(cdclevel0$wtdesire, 0.25)
quantile(cdclevel0$wtdesire, 0.75)

median(cdclevel1$wtdesire)
quantile(cdclevel1$wtdesire, 0.25)
quantile(cdclevel1$wtdesire, 0.75)

median(cdclevel0$loss)
quantile(cdclevel0$loss, 0.25)
quantile(cdclevel0$loss, 0.75)

median(cdclevel1$loss)
quantile(cdclevel1$loss, 0.25)
quantile(cdclevel1$loss, 0.75)

sum(cdclevel0$gender == 0)
sum(cdclevel0$gender == 0)/count(cdclevel0)

sum(cdclevel1$gender == 0)
sum(cdclevel1$gender == 0)/count(cdclevel1)

sum(cdclevel0$hlthplan == 1)
sum(cdclevel0$hlthplan == 1)/count(cdclevel0)

sum(cdclevel1$hlthplan == 1)
sum(cdclevel1$hlthplan == 1)/count(cdclevel1)

sum(cdclevel0$smoke100 == 1)
sum(cdclevel0$smoke100 == 1)/count(cdclevel0)

sum(cdclevel1$smoke100 == 1)
sum(cdclevel1$smoke100 == 1)/count(cdclevel1)

sum(cdclevel0$genhlth == "excellent")
sum(cdclevel0$genhlth == "excellent")/count(cdclevel0)

sum(cdclevel1$genhlth == "excellent")
sum(cdclevel1$genhlth == "excellent")/count(cdclevel1)

sum(cdclevel0$genhlth == "very good")
sum(cdclevel0$genhlth == "very good")/count(cdclevel0)

sum(cdclevel1$genhlth == "very good")
sum(cdclevel1$genhlth == "very good")/count(cdclevel1)

sum(cdclevel0$genhlth == "good")
sum(cdclevel0$genhlth == "good")/count(cdclevel0)

sum(cdclevel1$genhlth == "good")
sum(cdclevel1$genhlth == "good")/count(cdclevel1)

sum(cdclevel0$genhlth == "fair")
sum(cdclevel0$genhlth == "fair")/count(cdclevel0)

sum(cdclevel1$genhlth == "fair")
sum(cdclevel1$genhlth == "fair")/count(cdclevel1)

sum(cdclevel0$genhlth == "poor")
sum(cdclevel0$genhlth == "poor")/count(cdclevel0)

sum(cdclevel1$genhlth == "poor")
sum(cdclevel1$genhlth == "poor")/count(cdclevel1)
```




#Create empirical logit plots for the first research question
```{r}
emplogit(cdcEdit$age, cdcEdit$exerany, main = "Figure 1", xlab = "Age", ylab = "Log Odds of Exercising Regularly")

emplogit((cdcEdit$loss), (cdcEdit$exerany), main = "Figure 2", xlab = "Weight Loss Desired", ylab = "Log Odds of Exercising Regularly")
```
#Create empirical logit plots for the second research question
```{r}
emplogit(cdc$age, cdc$exerany, main = "Figure 1", xlab = "Age", ylab = "Log Odds of Exercising Regularly")

emplogit(abs(cdc$loss), (cdc$exerany), main = "Figure 2", xlab = "Weight Loss Desired", ylab = "Log Odds of Exercising Regularly")

emplogit(cdc$height, cdc$exerany, main = "Figure 1", xlab = "Height", ylab = "Log Odds of Exercising Regularly")

emplogit(cdc$weight, cdc$exerany, main = "Figure 1", xlab = "Weight", ylab = "Log Odds of Exercising Regularly")

emplogit((cdc$wtdesire), cdc$exerany, main = "Figure 1", xlab = "Weight Desire", ylab = "Log Odds of Exercising Regularly")

```


#Mosaic plots for first research question
```{r}
m1 <- table(cdcEdit$genhlth, cdcEdit$exerany)
dimnames(m1)[[1]] <- c("Excellent", "Very Good","Good", "Fair", "Poor")
dimnames(m1)[[2]] <- c("No", "Yes")

mosaicplot(m1, main = "Figure 4", xlab = "General Health", ylab = "Excercised In The Last Month")

m2 <- table(cdcEdit$hlthplan, cdcEdit$exerany)
dimnames(m2)[[1]] <- c("No", "Yes")
dimnames(m2)[[2]] <- c("No", "Yes")

mosaicplot(m2, main = "Figure 5", xlab = "Health Coverage", ylab = "Excercised In The Last Month")
```
#Research question for second research question
```{r}
m1 <- table(cdc$genhlth, cdc$exerany)
dimnames(m1)[[1]] <- c("Excellent", "Very Good","Good", "Fair", "Poor")
dimnames(m1)[[2]] <- c("No", "Yes")

mosaicplot(m1, main = "Figure 4", xlab = "General Health", ylab = "Excercised In The Last Month")

m2 <- table(cdc$hlthplan, cdc$exerany)
dimnames(m2)[[1]] <- c("No", "Yes")
dimnames(m2)[[2]] <- c("No", "Yes")

mosaicplot(m2, main = "Figure 5", xlab = "Health Coverage", ylab = "Excercised In The Last Month")

m3 <- table(cdc$smoke100, cdc$exerany)
dimnames(m3)[[1]] <- c("No", "Yes")
dimnames(m3)[[2]] <- c("No", "Yes")

mosaicplot(m3, main = "Figure 5", xlab = "Smoked Atleast 100 Cigarettes During Lifetime", ylab = "Excercised In The Last Month")

m4 <- table(cdc$gender, cdc$exerany)
dimnames(m4)[[2]] <- c("No", "Yes")
dimnames(m4)[[1]] <- c("Female", "Male")

mosaicplot(m4, main = "Figure 5", xlab = "Gender", ylab = "Excercised In The Last Month")
```
#Create correlation plot for 2nd research question
```{r}
cdc$absloss <- abs(cdc$loss)
cdcTitle <- data.frame("Height"= rep(NA,20000), "Weight"= rep(NA,20000), "Desired Weight"= rep(NA,20000), "Age"= rep(NA,20000), "Desired Weight Loss"= rep(NA,20000), "Absolute Value Desired Weight Loss"= rep(NA,20000))
cdcTitle$Height <- cdc$height
cdcTitle$Weight <- cdc$weight
cdcTitle$Desired.Weight <- cdc$wtdesire
cdcTitle$Age <- cdc$age
cdcTitle$Desired.Weight.Loss <- cdc$loss
cdcTitle$Abs.Weight.Loss <- cdc$absloss

corrplot(cor(cdc[,c("height","weight", "wtdesire", "age", "absloss")]),type = "upper",method = "num", title= "Figure 18", tl.cex = 0.7, mar = c(0,0,1,0))
corrplot(cor(cdcTitle[,c("Height","Weight", "Desired.Weight", "Age", "Abs.Weight.Loss")]),type = "upper",method = "num", tl.cex = 0.7, mar = c(0,0,1,0), tl.col = "black", col = "black")
```

#fit model for first research question
```{r}
glm1 <- glm(exerany ~ loss + age + genhlth + hlthplan, data = cdcEdit, family = "binomial")
summary <- round((summary(glm1)$coefficients),3)
knitr::kable(summary)
print(xtable(summary, type = "latex"))
cooks.distance(glm1)[which.max(cooks.distance(glm1))]
which.max(cooks.distance(glm1))>(4/2000)
plot(cooks.distance(glm1))

#predictions <- predict(glm1, newdata = cdcEdit, type = "response")

#fit <- glm(cdcEdit$exerany ~ predictions, #family = binomial)
#data.frame(x = predictions, residuals = qresid(fit)) %>%
#  ggplot(aes(x = predictions, y = residuals)) +
#  geom_point() +
#  theme_bw() + labs(x = "Predicted Probability", y = "Quantized Residuals") + geom_smooth(color = "grey")
```

#Find VIF for explanatory variables
```{r}
cdctryagain <- cdc[,c(1,2,3,4,5,6,7,8,9)]
glm7 <- glm(exerany ~., data = cdctryagain, family = binomial)
vif(glm7)
```

#Generate predictions for second model
```{r}
cdcvalidate <- cdc[,c(1,2,3,5,6,8,10)] #remove wtdesire, gender and smoking
#Creating Training and Test Data for all variables
n <- nrow(cdcvalidate)
set.seed(32)
# Sample the rows
rowsTrain <- sample(1:n, n*.8, replace = FALSE)
# Create the new training data
train <- cdcvalidate[rowsTrain, ]
# Create the new test data
test <- cdcvalidate[-rowsTrain, ]
true1 <- which(test$exerany == 1)
true0 <- which(test$exerany == 0)
ntrue1 <- length(true1)
ntrue0 <- length(true0)

storage <- data.frame("p"= rep(NA,n-length(rowsTrain)), "actual" = test$exerany)
glm2 <- glm(exerany ~., data = train, family = binomial)
storage$p <- predict(glm2, newdata = test, type = "response")

summary2 <- round((summary(glm2)$coefficients),3)
knitr::kable(summary2)
print(xtable(summary2, type = "latex"))
```

#Cook's distance for second model
```{r}
cooks.distance(glm2)[which.max(cooks.distance(glm2))]
plot(cooks.distance(glm2))
```
#Quantile residual plot
```{r}
fittest <- glm(test$exerany ~ storage$p, family = binomial)
data.frame(x = storage$p, residuals = qresid(fittest)) %>%
  ggplot(aes(x = storage$p, y = residuals)) +
  geom_point() +
  theme_bw() + labs(x = "Predicted Probability", y = "Quantized Residuals") + geom_smooth(color = "grey")
```


#ROC curve
```{r}
par(pty = "s")
roc.info <- roc(test$exerany, storage$p, plot = TRUE, legacy.axes = FALSE, auc=TRUE, ylab = "True Positive Rate (Sensitivity)", xlab = "False Positive Rate (1-Specificity)", xlim = (c(0.5,-.50)), ylim = (c(0,1)))
roc.df <- data.frame(Threshold = roc.info$thresholds, Sensitivity=roc.info$sensitivities, Specificity = (roc.info$specificities))
roc.df$Sum <- roc.df$Sensitivity + roc.df$Specificity
index <- which.max(roc.df$Sum)
table2 <- roc.df[index,]
table2$AUC <- roc.info$auc
table2 <- round(table2 , 4)
knitr::kable(table2)
print(xtable(table2, type = "latex"))

plot.roc(
  roc.info, 
  print.auc = TRUE, 
  auc.polygon = TRUE, 
  grid = TRUE, 
  grid.col = "lightgray", 
  xlab = "False Positive Rate (1-Specificity)", 
  ylab = "True Positive Rate (Sensitivity)", 
  #xlim = (c(0,-1))
)
```


